---
title: "Bruce Campbell ST-617 Discussion Group 7"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'pdf')
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(prompt=FALSE)
knitr::opts_chunk$set(fig.height=5)
knitr::opts_chunk$set(fig.width=7)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=TRUE)
```

`r date()`



```{r }
rm(list = ls())
set.seed(7)

setwd("C:/st-617/")

## use read.csv2 since data fields are delimited by semicolons (;)

cls=c(rep("numeric",11),"integer")

wine.data <- read.csv2("FinalProject/winequality-red.csv")

#Specifying the data type via colClasses did not work - so we sapply as.numeric to convert the factor data to numeric.  We may want to convert some of the variables to factors later - but for visualization of the raw features we should use numeric type. 
wine.data[, c(1:12)] <- sapply(wine.data[, c(1:12)], as.numeric)
```


```{r}

for ( i in 1:(ncol(wine.data)-1))
{
  feature = wine.data[,i]
  featureName = names(wine.data)[i]
  hist(feature, main = featureName,freq = FALSE,50)
}
  i=ncol(wine.data)
  feature = wine.data[,i]
  featureName = names(wine.data)[i]
  hist(feature, main = featureName,freq = FALSE)
```

#### _fixed.acidity_
Slightly bimodal 

#### _volatile.acidity_     
Fat tail.  Looks Poisson

#### _citric.acid_          


#### _residual.sugar_      
Bimodal, fat tail to the right

#### _chlorides_            
Fat tail to the right.  Looks lognormal

#### _free.sulfur.dioxide_


#### _total.sulfur.dioxide_

#### _density_             

#### _pH_  
Looks Normal

#### _sulphates_  
Fat tail to the right - looks lognormal

#### _alcohol_   
Slightly bimodal

#### _quality_  
The majority of the wines have a quality factor of 5 ,6




