---
title: "Bruce Campbell ST-617 Discussion Problem #2"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'pdf')
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(prompt=FALSE)
knitr::opts_chunk$set(fig.height=11)
knitr::opts_chunk$set(fig.width=8)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=TRUE)
```

`r date()`

#Chapter 3
##Problem 3
```
Suppose we have a data set with five predictors, X1 =GPA, X2 = IQ,
X3 = Gender (1 for Female and 0 for Male), X4 = Interaction between
GPA and IQ, and X5 = Interaction between GPA and Gender. The
response is starting salary after graduation (in thousands of dollars).
Suppose we use least squares to fit the model, and get 
```
$$\beta_0 = 50, \beta_1 =20, \beta_2 = 0.07, \beta_3 = 35, \beta_4 = 0.01, \beta_5 = -10$$

The linear model is then

$$ Salary = 50 + 20 * GPA + 0.07 * IQ + 35 * Gender + 0.01 * (GPA * IQ) -10 * (GPA * Gender)$$


```{r}
rm(list = ls())
GPA <- (0.0:0.2:4.0)
```

## a) Which answer is correct, and why?

```
i. For a fixed value of IQ and GPA, males earn more on average
than females.
ii. For a fixed value of IQ and GPA, females earn more on
average than males.
iii. For a fixed value of IQ and GPA, males earn more on average
than females provided that the GPA is high enough.
iv. For a fixed value of IQ and GPA, females earn more on
average than males provided that the GPA is high enough.
```

Let's rewrite our regression equation, gathering the gender term

$$Salary = Gender * \big(  35 + -10 GPA )  \big)  + \big( 50 + 20 * GPA + 0.07 * IQ + 0.01 * (GPA * IQ) \big)$$

Note that for all the values of IQ and GPA the term on the right is positive, that gender is a binary predictor
and that range of GPA is $[0,4]$.  We see that when Gender=1 (female) and $GPA<3.5$ the contribution from gender 
to salary is positive.  Therefore the correct statement is i).  


## b) ```Predict the salary of a female with IQ of 110 and a GPA of 4.0.```

```{r}
GPA <- 4.0
IQ <- 110
Gender <-1
Salary <- 50 + 20 * GPA + 0.07 * IQ + 35 * Gender + 0.01 * (GPA * IQ) -10 * (GPA * Gender)
```

The predicted salary is `r Salary`

## c) ```
True or false: Since the coefficient for the GPA/IQ interaction
term is very small, there is very little evidence of an interaction
effect. Justify your answer.
```

This is a true statement. To validate it we need to consider the scale of the perdictors, and ideally we'd like to see p-values as well. 

$$range(GPA * IQ) =[ min(GPA) * min(IQ),max(GPA) * max(IQ) ] \approx [0,800]$$

More realistically, the range is probably $[50,800]$, at this level all other variables held fixed 
for every unit increase in GPA*IQ we expect a 0.01 increase in salary (in units of $1,000) - or an additional $10.
At the midpoints of the range of IQ * GPA (400) we have an increase of o fonly 4000. Given that the $\beta_0 = 50$, and
the other coefficients and perdictor ranges, we are lead to the conclusion that this interaction term is not significant. 




