---
title: "Bruce Campbell ST-617 Homework 2"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'pdf')
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(prompt=FALSE)
knitr::opts_chunk$set(fig.height=11)
knitr::opts_chunk$set(fig.width=8)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=TRUE)
```

`r date()`

#Chapter 4

##Problem 7
Suppose that we wish to predict whether a given stock will issue a
dividend this year ("Yes" or "No") based on X, last year's percent
profit.We examine a large number of companies and discover that the
mean value of X for companies that issued a dividend was $\overline{X} = 10$,
while the mean for those that didn't was $\overline{x} = 0$. In addition, the
variance of X for these two sets of companies was $\hat{\sigma^2} = 36$. Finally,
80% of companies issued dividends. Assuming that X follows a normal
distribution, predict the probability that a company will issue
a dividend this year given that its percentage profit was X = 4 last
year.


Let's denote our 2 classes 0,1 where 0 indicates no dividend and 1 indicates a dividend. Then 
we will need to calculate the posterior probability $P(Y=1 | X=4)$.  We are given all of the information we need to 
do this.  The prior probabilities are $\pi_1=0.8$ and $\pi_0=0.2$ and the likelihood of each class is given 
by $N(\mu_1,\hat{\sigma})(x)$ where $\mu_1=10$ and $N(\mu_0,\hat{\sigma})$ where $\mu_0=0$ and $\hat{\sigma} = 36$ in both cases.
$N(\mu,\sigma)(x)= \frac{1}{\sqrt{2 \pi} \sigma} e^{\frac{-1}{2 \sigma^2} (x-\mu)^2}$ is the normal distribution with mean $\mu$ and 
variance $\hat{\sigma^2}$.

Putting this all together into Bayes theorem

$$P(Y=1 | X=4) = \frac {N(\mu_1,\hat{\sigma})(4) \pi_1}{N(\mu_1,\hat{\sigma})(4) \pi_1 + N(\mu_0,\hat{\sigma})(4) \pi_0 }$$

```{r}

mu_1=.1
mu_0=0

pi_1=.8
pi_0=.2

sigma_sq=36
stdev_est=6

x=4

posterior_probability_of_dividend_given_x <- function(x,mu_1,mu_0,stdev_est, pi_1,pi_0)
{
  probability = ( dnorm(x, mean = mu_1, sd = stdev_est) * pi_1 ) /(  dnorm(x, mean = mu_1, sd = stdev_est) * pi_1  + dnorm(x, mean = mu_0, sd = stdev_est) * pi_0 )
  return(probability)
}
  
posterior_probability_of_dividend <- posterior_probability_of_dividend_given_x(x,mu_1,mu_0,stdev_est, pi_1,pi_0)

```

We have that the probability of a dividend in the event the $X=$ is ```r posterior_probability_of_dividend```.



