---
title: "Bruce Campbell ST-617 Homework 2"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'pdf')
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(prompt=FALSE)
knitr::opts_chunk$set(fig.height=11)
knitr::opts_chunk$set(fig.width=8)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)
```

`r date()`

#Chapter 5

##Problem 5
In Chapter 4, we used logistic regression to predict the probability of
default using income and balance on the Default data set. We will
now estimate the test error of this logistic regression model using the
validation set approach. Do not forget to set a random seed before
beginning your analysis.

## a) 
Fit a logistic regression model that uses income and balance to
predict default.

```{r }
rm(list = ls())
library(ISLR)
attach(Default)
glm.fit = glm(default~income+balance,data=Default, family =binomial)
summary(glm.fit)
```

### b) 
Using the validation set appbroach, estimate the test error of this
model. In order to do this, you must perform the following steps:
i. Split the sample set into a training set and a validation set.
ii. Fit a multiple logistic regression model using only the training
observations.
iii. Obtain a prediction of default status for each individual in
the validation set by computing the posterior probability of
default for that individual, and classifying the individual to
the default category if the posterior probability is greater
than 0.5.
iv. Compute the validation set error, which is the fraction of
the observations in the validation set that are misclassified.

```{r}
set.seed(7)

train=sample(nrow(Default), floor(nrow(Default)* 2/3))
DF<-Default
DFTrain <-DF[train,]
DFTest <-DF[-train,]

glm.fit = glm(default~income+balance,data=DFTrain, family =binomial)
summary(glm.fit)

glm.probs =predict (glm.fit ,DFTest,type ="response")

contrasts(DFTest$default)
glm.pred=rep ("No " ,nrow(DFTest))
glm.pred[glm.probs >.5]=" Yes"
TB<-table(glm.pred ,DFTest$default )
ACC_Validation = (TB[2]+TB[3])/ length(DFTest$default)

modelsDF <-data.frame(iteration=1,Accuracy=1-ACC_Validation)

```

### c)
Repeat the process in (b) three times, using three different splits
of the observations into a training set and a validation set. Comment
on the results obtained.
```{r}
attach(Default)
modelsDF <- data.frame(iteration=numeric(), Accuracy=numeric())

for(i in 1: 3)
{
  train=sample(nrow(Default), floor(nrow(Default)* 2/3))
  DF<-Default
  DFTrain <-DF[train,]
  DFTest <-DF[-train,]
  
  glm.fit = glm(default~income+balance,data=DFTrain, family =binomial)
  summary(glm.fit)
  
  glm.probs =predict (glm.fit ,DFTest,type ="response")

  glm.pred=rep ("No " ,nrow(DFTest))
  glm.pred[glm.probs >.5]=" Yes"
  TB<-table(glm.pred ,DFTest$default )
  ACC_Validation = (TB[2]+TB[3])/ length(DFTest$default)
  modelsDF <-rbind(modelsDF,data.frame(iteration=i,Accuracy=1-ACC_Validation))
}
library(pander)
pander(modelsDF)
```

The validation test set error rates are all very similar.  This indicated the model is stable with respect to the random split into test and training
sets and that the validation approach may be vaible in this instance althugh we'd probably want to do more iterations to confirm this.

### d) 
Now consider a logistic regression model that predicts the probability
of default using income, balance, and a dummy variable
for student. Estimate the test error for this model using the validation
set approach. Comment on whether or not including a
dummy variable for student leads to a reduction in the test error
rate.
```{r}
attach(Default)
modelsDFAug <- data.frame(iteration=numeric(), Accuracy=numeric())

for(i in 1: 3)
{
  train=sample(nrow(Default), floor(nrow(Default)* 2/3))
  DF<-Default
  DFTrain <-DF[train,]
  DFTest <-DF[-train,]
  
  glm.fit = glm(default~income+balance+student,data=DFTrain, family =binomial)
  summary(glm.fit)
  
  glm.probs =predict (glm.fit ,DFTest,type ="response")

  glm.pred=rep ("No " ,nrow(DFTest))
  glm.pred[glm.probs >.5]=" Yes"
  TB<-table(glm.pred ,DFTest$default )
  ACC_Validation = (TB[2]+TB[3])/ length(DFTest$default)
  modelsDFAug <-rbind(modelsDFAug,data.frame(iteration=i,Accuracy=1-ACC_Validation))
}
library(pander)
pander(modelsDFAug)
```

Including the student status as a predictor did not appear to change the validation set error rates - the change is the number of errors for each of the three runs is : 
```{r}
diff<- (modelsDFAug$Accuracy - modelsDF$Accuracy)*nrow(DFTest)
pander(diff)
```

To make a more precise statement we'd run more iterations and compare the errors using a statistical test such as a t-test. 

