---
title: "Bruce Campbell ST-617 Homework 3"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'pdf')
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(prompt=FALSE)
knitr::opts_chunk$set(fig.height=5)
knitr::opts_chunk$set(fig.width=7)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)
```

`r date()`

#Chapter 7

##Problem 10
This question relates to the College data set.

### a) 
Split the data into a training set and a test set. Using out-of-state
tuition as the response and the other variables as the predictors,
perform forward stepwise selection on the training set in order
to identify a satisfactory model that uses just a subset of the
predictors.

```{r}
rm(list = ls())
library(ISLR)
attach(College)

train=sample(nrow(College), floor(nrow(College)* 2/3))
DF<-College
DFTrain <-DF[train,]
DFTest <-DF[-train,]

library(pander)
pander(names(DF))

library(leaps)
regfit.full<-regsubsets(Outstate~.,data = DFTrain,method = "forward",nvmax = 18)

reg.summary <- summary(regfit.full)

mse_v <- reg.summary$rss / nrow(DF)

plot(mse_v) 
title(c("MSE versus model size for forward subset selection algorithm on training set.","Forward SSS"))

plot(regfit.full ,scale ="bic")
title("$BIC$ Forward SSS")

model_fss8 <- coef(regfit.full,8)
pander(names(model_fss8))
```

### b) 
Fit a GAM on the training data, using out-of-state tuition as
the response and the features selected in the previous step as
the predictors. Plot the results, and explain your findings.

```{r,fig.height=4.7, fig.width=6.7}
library(gam)

gam.fit=gam(Outstate~Private+s(Room.Board,4)+s(Personal,4)+s(Terminal,4)+s(S.F.Ratio,4)+s(perc.alumni)+s(Expend,4)+s(Grad.Rate,4) ,data=DFTrain)

plot(gam.fit, se=TRUE ,col ="blue ")

```

We have used the gam function to fit a univariate smmothing spline with 4 degrees of freedon to each perdictor.  The plots show the univariate fits.
One of the predictors selected by the forward SSS algorithm is a factor and is not fit to a smoothing spline. 
There is strong evidence of non linear relationships in the data.  The variables ${Personal, S.F.Ratio, perc.alumni, expend}$ show this particularly. 

### c) 
Evaluate the model obtained on the test set, and explain the
results obtained.

```{r}
preds=predict (gam.fit,newdata =DFTest)

RSS <- sum((preds-DFTest$Outstate)^2)
TSS <- sum((DFTest$Outstate - mean(DFTest$Outstate))^2)
RS2_Test <- 1- (RSS/TSS)

plot(DFTest$Outstate,DFTest$Outstate-preds)
title(c("Residual plot for test set",sprintf("R-Squared = %f",RS2_Test)))

preds=predict (gam.fit,newdata =DFTrain)
RSS <- sum((preds-DFTrain$Outstate)^2)
TSS <- sum((DFTrain$Outstate - mean(DFTrain$Outstate))^2)
RS2_Train <- 1- (RSS/TSS)
```

We see no significant trend in the residual plot, indicating that there is no unaccounted for non-linear relationships in the model.
We also see that the training and test set  $R^2$ statistic indicate a resonable fit.  As expected the test $R^2$ statistic is below the
training $R^2$ value. 

### d) 
For which variables, if any, is there evidence of a non-linear
relationship with the response?

The variables ${Personal, S.F.Ratio, perc.alumni, expend}$ particularly show a non -linear relationship with the response.
