---
title: "Bruce Campbell ST-617 Homework 2"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'pdf')
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(prompt=FALSE)
knitr::opts_chunk$set(fig.height=5)
knitr::opts_chunk$set(fig.width=4)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)
```

`r date()`

#Chapter 5

##Problem 9
We will now consider the Boston housing data set, from the MASS
library.
### a) 
Based on this data set, provide an estimate for the population
mean of medv. Call this estimate $\hat{\mu}$
```{r}
library(MASS)
attach(Boston)
mu_hat <- mean(Boston$medv)
```
$\hat{\mu}=$ ```r mu_hat```

### b) 
Provide an estimate of the standard error of $\hat{\mu}$. Interpret this
result.

Hint: We can compute the standard error of the sample mean by
dividing the sample standard deviation by the square root of the
number of observations.

```{r}
se_sm <- sd(Boston$medv)/nrow(Boston)
```
 Our estimate of the standard error of the sample mean based on the standard deviation
 is ```r se_sm```

### c)
Now estimate the standard error of $\hat{\mu}$ using the bootstrap. How
does this compare to your answer from (b)?
```{r}
library(boot)
alpha.fn <-function(data,index)
{
  D=data[index,]
  result <- mean(D$medv)
  return(result)
}

se_bootstrap <-boot(Boston,alpha.fn,100)

se_bootstrap
```

### d) 
Based on your bootstrap estimate from (c), provide a 95% confidence
interval for the mean of medv. Compare it to the results
obtained using t.test(Boston$medv).

Hint: You can approximate a 95% confidence interval using the
formula $[\hat{\mu} - 2SE(\hat{\mu}), \hat{\mu} + 2SE(\hat{\mu})]$.

```{r}
left_ci <-  mu_hat - 2 * sd(se_bootstrap$t)
right_ci<- mu_hat + 2 * sd(se_bootstrap$t)
```


Our 95% CI as estimated from the boostrap calculation is
[```r left_ci``` , ```r right_ci```] 

The results of the t-test are comparable to the bootstrap estimate, but the one sample t-test provides a smaller estimate of the 95% CI.

```{r}
t.test(Boston$medv)
```


### e) 
Based on this data set, provide an estimate, med, for the median
value of medv in the population.

We can use the sample median for an estimate of the population median which is :

```{r}
median(Boston$medv)
```

### f)
We now would like to estimate the standard error of $\hat{\mu_med}$. Unfortunately,
there is no simple formula for computing the standard
error of the median. Instead, estimate the standard error of the
median using the bootstrap. Comment on your findings.
```{r}
alpha.fn <-function(data,index)
{
  D=data[index,]
  result <- median(D$medv)
  return(result)
}
se_median_bootstrap <-boot(Boston,alpha.fn,100)

se_median_bootstrap
```

The estimate of standard error of the median as calculated by the bootstrap is 
```{r}
sd(se_median_bootstrap$t)
```


```{r}
left_ci <-  mu_hat - 2 * sd(se_median_bootstrap$t)
right_ci<- mu_hat + 2 * sd(se_median_bootstrap$t)
```

Our 95% CI as estimated from the boostrap calculation is
[```r left_ci``` , ```r right_ci```] 

We're not sure if this estimate is valid for the median so let's calculate the 5th and 95th quantile from the bootstrap sample

```{r}
left_ci <-  quantile(se_median_bootstrap$t,0.05)
right_ci<-  quantile(se_median_bootstrap$t,0.95)
```

[```r left_ci``` , ```r right_ci```] 

This is a much tighter CI, so we wonder if the estimate above applied to the median.

### g) 
Based on this data set, provide an estimate for the tenth percentile
of medv in Boston suburbs. Call this quantity $\hat{\mu_0.1}$. (You
can use the quantile() function.)

The sample tenth percentile of medv is
```{r}
quantile(Boston$medv,.1)
```

###h) 
Use the bootstrap to estimate the standard error of $\hat{\mu_0.1}$. Comment
on your findings.

```{r}
alpha.fn <-function(data,index)
{
  D=data[index,]
  result <- quantile(D$medv, 0.1)
  return(result)
}
se_quantile_bootstrap <-boot(Boston,alpha.fn,100)

se_quantile_bootstrap
```

The estimate of standard error of the tenth percentile as calculated by the bootstrap is 
```{r}
sd(se_quantile_bootstrap$t)
```
