---
title: "Bruce Campbell ST-617 HW 1"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'pdf')
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(prompt=FALSE)
knitr::opts_chunk$set(fig.height=11)
knitr::opts_chunk$set(fig.width=8)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=TRUE)
```

`r date()`

#Chapter 2

##Problem 10

### a)
There are 506 observations in the Boston data set which contain a variety of variables porported to affect housing values in the suburbs of Boston. 

This data frame contains the following columns:

* crim : per capita crime rate by town.

* zn :proportion of residential land zoned for lots over 25,000 sq.ft.

* indus : proportion of non-retail business acres per town.

* chas :Charles River dummy variable (= 1 if tract bounds river; 0 otherwise).

* nox :nitrogen oxides concentration (parts per 10 million).

* rm :average number of rooms per dwelling.

* age : proportion of owner-occupied units built prior to 1940.

* dis : weighted mean of distances to five Boston employment centres.

* rad : index of accessibility to radial highways.

* tax : full-value property-tax rate per \$10,000.

* ptratio : pupil-teacher ratio by town.

* black : 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town.

* lstat : lower status of the population (percent).

* medv : median value of owner-occupied homes in \$1000s

### b) 
```{r}
library(MASS)
#library(ggplot2)
#require(GGally)
library(pander)
DF <-Boston
```
![Scatterplot](image.pdf)

We note 
* several bimodal variables : indus,rad,tax
* some linear relationships :  (dis,age) (rm,mdev)
* a few mildly non linear relationships (nox,age) (nox,dis)

### c) 
chas has a higher crime rate at a value of 0

`r plot(DF$chas,DF$crim)`

certain values of nox (nox >.57 ) are associated with a higher crime rate 

`r plot(DF$nox,DF$crim)`

older towns have a higher crime rate

`r plot(DF$age,DF$crim)`

Additionally there is a spike in crime at a indus value of 19

`r plot(DF$indus,DF$crim)`

### d)
The crime rate variable has a long tail. 

`r hist(DF$crim,50)`

There are 11 obersevations with a value higher than 25.

`r p <- DF[DF$crim>25,]; `

The tax rate appears bimodal with 137 towns having a rate above 650 and the rest below 450

`r hist(DF$tax,50)`

There are some towns with a low putpil teacher ratio. We note that the tax rate for these towns does not fall in the upper bracket. 

We use the summary function to look at the ranges of the variables
```{r}
S<-summary(DF)
pander(S)
```
### e) 

```{r}
countBoundingCharles <- nrow(DF[DF$chas==1,])
```
There are `r countBoundingCharles` towns identified as bordering the Charles.

### f) 
From the summary table above we see that the median pupil to teacher ration is 19.05.

### g)
```{r}
lowMed <- DF[which.min(DF$medv), ]
pander(lowMed)
```
The town with the lowest median value of owner occupied houses has a high crime rate as indicated by the box plot below.

```{r,dpi=1200,tidy=TRUE,prompt=FALSE,echo=FALSE,fig.height=5, fig.width=5,dev="pdf"}
boxplot(DF$crim)
points(lowMed$crim, col='red', pch='*',cex=3)
title("Crime Rate with min(meddev) indicated in red")
```

Nox is elevated for this town

```{r,dpi=1200,tidy=TRUE,prompt=FALSE,echo=FALSE,fig.height=5, fig.width=5,dev="pdf"}
boxplot(DF$nox)
points(lowMed$nox, col='red', pch='*',cex=3)
title("Nos with min(meddev) indicated in red")
```

Indus is elevated for this town

```{r,dpi=1200,tidy=TRUE,prompt=FALSE,echo=FALSE,fig.height=5, fig.width=5,dev="pdf"}
boxplot(DF$indus)
points(lowMed$indus, col='red', pch='*',cex=3)
title("Indus with min(meddev) indicated in red")
```

Black is elevated for this town
```{r,dpi=1200,tidy=TRUE,prompt=FALSE,echo=FALSE,fig.height=5, fig.width=5,dev="pdf"}
boxplot(DF$black)
points(lowMed$black, col='red', pch='*',cex=3)
title("Black with min(meddev) indicated in red")
```


### h) 

```{r}
sevenRooms<- DF[DF$rm>7,]

eightRooms<- DF[DF$rm>8,]
```

There are `r nrow(sevenRooms)` dwellings with more than seven rooms, and there are `r nrow(eightRooms)` with more than eight rooms.  

```{r}
pander(eightRooms)

```






