---
title: "ST 502 HW 3 Chapter 7 Problem 65"
author: "Bruce Campbell"
date: "January 29, 2017"
output: pdf_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE} 
rm(list = ls())
knitr::opts_chunk$set(dev = 'png')
knitr::opts_chunk$set(fontsize=13)
knitr::opts_chunk$set(dpi=600)
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(prompt=FALSE)
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(fig.height=6)
knitr::opts_chunk$set(fig.width=6)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)
```

(a) Suppose that truly our population follows a $N(\mu = 3, \sigma^2 = 4)$ distribution. We
want to inspect how well certain CIs for $\mu$ will work. Generate a random sample
of size n=5 from the population. For that sample, save the MOM (also the MLE
here) estimate of $\mu$ , $\hat{\mu} = \bar{Y}$, and the unbiased estimate of $\sigma^2$ 

$$  S^2 = \frac{1}{n-1} \sum\limits_{i=1}^{n} (Y_i - \bar{Y})^2 $$

```{r }
library(pander)
library(plyr)
library(dplyr)
library(readr)

mu <- 3
sigma <- 2
sampleSize <- 5
sample5 <- rnorm(5,mu,sigma)

muHat <- mean(sample5)

S2 <- var(sample5)

# Make sure R is using unbiased estimator 
#     sum(sample5^2 - muHat^2)/(sampleSize-1)

```


(b) Create and save a 95% CI for $\mu$ using only the data you have (i.e. you must use
the t-distribution).

Since our sample is from a normal distribution we have a pivotal quantity from which we can make a probaility statement that will be the basis of our confidence interval.

$$ \frac{ \sqrt{n}  (\bar(Y) - \mu  }{ S }  \sim t_{n-1}  $$

Letting $t(\frac{\alpha}{2})$ be the $frac{\alpha}{2}$ quantile and noting that $t(\frac{\alpha}{2}) = t(1 - \frac{\alpha}{2})$ by symmetry, we have   

$$  P(  \bar{Y} - \frac{S}{\sqrt{n}} t(\frac{\alpha}{2})  \leq \mu \leq \bar{Y} + \frac{S}{\sqrt{n}} t(\frac{\alpha}{2}) )  = 1-\alpha $$

```{r}

S <- sqrt(S2)
alpha <- 0.05
talpha <- qt(1-alpha/2,df = sampleSize-1)

CI <- c(muHat - S/sqrt(sampleSize) * talpha, muHat + S/sqrt(sampleSize) * talpha)

pander(data.frame(CI), caption = "95% CI")

```


(c) We will also create a bootstrap percentile interval for $\mu$. For the RS of size five,
use your observed MOM estimators to bootstrap $B=500$ data sets (parametric
bootstrap). Save the mean for each bootstrapped data set.


(d) Create a 95% CI for $\mu$ using the bootstrap percentile method.

````{r}
sampleSize <- 5
numSamples <- 50
#Define a function to run the simulations
run.CI <- function(sampleSize, numSamples)
{
  CIs <- data.frame(matrix(ncol = 2,nrow = numSamples))
  colnames(CIs) <- c("Xbar","s2")
  for (i in 1 : numSamples)
  {
    srs <- rnorm(sampleSize,mean = muHat,sd = sqrt(S2))
    CIs[i,]$Xbar <- mean(srs)
    CIs[i,]$s2 <- var(srs)
  }
  return(CIs)
}

df <-run.CI(sampleSize,numSamples)

rightCI <-quantile(df$Xbar,0.975)
leftCI <- quantile(df$Xbar,0.025)

pander(data.frame(left=as.numeric(leftCI),right=as.numeric(rightCI)),caption = " 95% CI from bootstrap quantiles")
```

(e) Repeat the above steps a total of N=5000 times.


(g) At the bottom of your file, report the proportion of intervals that contained the
true value of $\mu= 3$  for each sample size and method.

```{r}
sampleSize <- 5
numCIs <- 500

run.BS <- function(sampleSize, numCIs)
{
  BSCIs <- data.frame(matrix(ncol = 4,nrow = numCIs))
  colnames(BSCIs) <- c("left","right","containsP","CILength")
    
  for(i in 1:numCIs)
  {
    df <-run.CI(sampleSize,numSamples)
    BSCIs[i,]$right <-quantile(df$Xbar,0.975)
    BSCIs[i,]$left <- quantile(df$Xbar,0.025)
    BSCIs[i,]$containsP <- mu >= BSCIs[i,]$left & mu <= BSCIs[i,]$right
    BSCIs[i,]$CILength <-BSCIs[i,]$right -  BSCIs[i,]$left
  }
  return(BSCIs)
  
}

df <- run.BS(sampleSize = sampleSize, numCIs = numCIs)

proportion_n5_B500_quantile <-sum(df$containsP)/numCIs
meanCILength_n5_B500_quantile <- mean(df$CILength)

print(proportion_n5_B500_quantile)
print(meanCILength_n5_B500_quantile)
```


(f) Repeat all of the above for n = 20 and n = 100.
```{r}
sampleSize <- 20


df <- run.BS(sampleSize = sampleSize, numCIs = numCIs)

proportion_n20_B500_quantile <-sum(df$containsP)/numCIs
meanCILength_n20_B500_quantile <- mean(df$CILength)

print(proportion_n20_B500_quantile)
print(meanCILength_n20_B500_quantile)
```

n=100

```{r}
sampleSize <- 100


df <- run.BS(sampleSize = sampleSize, numCIs = numCIs)


proportion_n100_B500_quantile <-sum(df$containsP)/numCIs
meanCILength_n100_B500_quantile <- mean(df$CILength)

print(proportion_n100_B500_quantile)
print(meanCILength_n100_B500_quantile)
```


(h) Report the mean CI length for each sample size and method.
```{r}
pander(data.frame(meanCILength_n5_B500_quantile=meanCILength_n5_B500_quantile,meanCILength_n20_B500_quantile=meanCILength_n20_B500_quantile,meanCILength_n100_B500_quantile=meanCILength_n100_B500_quantile), caption = "length")
```





